---
title: "compare methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{compare methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

Load packages

```{r setup}
library(tidyr)       # Tidy Messy Data
# library(knitr)       # A General-Purpose Package for Dynamic Report Generation in R
library(kableExtra)  # Construct Complex Table with 'kable' and Pipe Syntax

library(ggBA)
```

Set options

```{r options}
options(
  kable_styling_bootstrap_options = c("hover", "striped"),
  kable_styling_full_width = FALSE,
  kable_styling_position = "left"
)
```


## Data

Prepare data: create columns for every method of measuring temperature.

```{r dataprep}
tbl <- 
  temperature |> 
  pivot_wider(names_from = method, values_from = temperature)
```

## Compare methods (identity scale)

### Compute statistics

```{r simple_stats_table}
ba_stat(data = tbl, var1 = infrared, var2 = rectal) |> 
  pivot_wider(names_from = parameter, values_from = value) |> 
  kable() |> 
  kable_styling()
```

### Visualize comparison

```{r simple_baplot}
ba_plot(data = tbl, var1 = infrared, var2 = rectal)
```

## Compare methods (log scale)

When measurements span several orders of magnitude, or when proportional
differences are more meaningful than absolute differences, a log-scale
Bland-Altman analysis is appropriate.  On the log scale the "difference"
becomes `log(var1) - log(var2) = log(var1 / var2)`, and the bias
represents a systematic ratio between the two methods.

```{r log_stats_table}
ba_stat(data = tbl, var1 = infrared, var2 = rectal, transform = "log") |> 
  pivot_wider(names_from = parameter, values_from = value) |> 
  kable(digits = 4) |> 
  kable_styling()
```

The bias on the log scale can be back-transformed to a ratio:
`exp(bias)` gives the geometric mean ratio (infrared / rectal).

```{r log_bias_backtransform}
log_stats <- ba_stat(data = tbl, var1 = infrared, var2 = rectal, transform = "log")
log_bias  <- log_stats$value[log_stats$parameter == "bias"]
cat("Geometric mean ratio (infrared / rectal):", round(exp(log_bias), 4), "\n")
```

Visualize the log-scale Bland-Altman plot.  The y-axis shows
`log(infrared / rectal)` and the x-axis shows the mean of the log-transformed
values.

```{r log_baplot}
ba_plot(
  data      = tbl,
  var1      = infrared,
  var2      = rectal,
  transform = "log",
  xlab      = "Mean of log(temperature)",
  ylab      = "log(infrared / rectal)"
)
```

## Compare proportions (logit scale)

For measurements expressed as proportions (values in (0, 1)), a logit
transformation stabilises variance and yields differences on an
approximately Normal scale.

```{r logit_example}
set.seed(42)
prop_tbl <- data.frame(
  p1 = runif(60, 0.05, 0.95),
  p2 = runif(60, 0.05, 0.95)
)

ba_stat(data = prop_tbl, var1 = p1, var2 = p2, transform = "logit") |>
  pivot_wider(names_from = parameter, values_from = value) |>
  kable(digits = 3) |>
  kable_styling()
```

```{r logit_baplot}
ba_plot(
  data      = prop_tbl,
  var1      = p1,
  var2      = p2,
  transform = "logit",
  xlab      = "Mean of logit(proportion)",
  ylab      = "Difference on logit scale"
)
```
