---
title: "Compare methods across analysis scales"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compare methods across analysis scales}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

```{r setup}
library(dplyr)
library(knitr)
library(tidyr)
library(ggBA)
```

This vignette compares the same two measurement methods under three transformations:

1. `identity`: absolute differences on the original scale
1. `log`: ratio-focused agreement
1. `logit`: agreement for proportion-scale measurements

## Prepare paired data

```{r dataprep}
tbl <- temperature |>
  pivot_wider(names_from = method, values_from = temperature)
```

The `temperature` dataset is stored in long format (`method` + `temperature` values).
For pairwise comparisons, we first create one column per method (`infrared`, `rectal`).

## Identity scale (default)

Use this when absolute differences are directly interpretable.

```{r simple_stats_table}
ba_stat(data = tbl, var1 = infrared, var2 = rectal) |>
  pivot_wider(names_from = parameter, values_from = value) |>
  kable(digits = 3)
```

```{r simple_baplot}
ba_plot(data = tbl, var1 = infrared, var2 = rectal)
```

## Log scale

Use a log transform when proportional differences are more meaningful than
absolute ones. On this scale:

`log(var1) - log(var2) = log(var1 / var2)`

```{r log_stats_table}
log_stats <- ba_stat(data = tbl, var1 = infrared, var2 = rectal, transform = "log")

log_stats |>
  pivot_wider(names_from = parameter, values_from = value) |>
  kable(digits = 4)
```

The bias can be back-transformed to a ratio:
`exp(bias)` gives the geometric mean ratio (infrared / rectal).

```{r log_bias_backtransform}
log_bias <- log_stats |>
  filter(parameter == "bias") |>
  pull(value)
cat("Geometric mean ratio (infrared / rectal):", round(exp(log_bias), 4), "\n")
```

```{r log_baplot}
ba_plot(
  data      = tbl,
  var1      = infrared,
  var2      = rectal,
  transform = "log",
  xlab      = "Mean of log(temperature)",
  ylab      = "log(infrared / rectal)"
)
```

## Logit scale for proportions

For values restricted to `(0, 1)`, a logit transform often gives better
variance behavior and an approximately Normal difference scale.

```{r logit_example}
set.seed(42)
prop_tbl <- data.frame(
  p1 = runif(60, 0.05, 0.95),
  p2 = runif(60, 0.05, 0.95)
)

logit_stats <- ba_stat(data = prop_tbl, var1 = p1, var2 = p2, transform = "logit")

logit_stats |>
  pivot_wider(names_from = parameter, values_from = value) |>
  kable(digits = 3)
```

```{r logit_baplot}
ba_plot(
  data      = prop_tbl,
  var1      = p1,
  var2      = p2,
  transform = "logit",
  xlab      = "Mean of logit(proportion)",
  ylab      = "Difference on logit scale"
)
```
