---
title: "Getting started with ggBA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ggBA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What this guide covers

This short guide walks through a complete Bland-Altman workflow:

1. reshape long-format data into paired columns
1. compute agreement statistics
1. visualize agreement with one line of plotting code

```{r setup}
library(knitr)
library(tidyr)
library(ggBA)
```

## 1) Prepare paired measurements

`temperature` contains one row per animal, visit, and measurement method.  
To compare two methods directly, pivot to wide format.

```{r prep-data}
tbl <- temperature |>
  pivot_wider(names_from = method, values_from = temperature)

head(tbl, 6) |>
  kable(digits = 2)
```

## 2) Compute Bland-Altman statistics

```{r compute-stats}
stats_tbl <- ba_stat(tbl, infrared, rectal)
stats_tbl |>
  tidyr::pivot_wider(names_from = parameter, values_from = value) |>
  kable(digits = 3)
```

The most commonly interpreted quantities are:

1. `bias`: average difference between methods
1. `lloa` and `uloa`: lower and upper limits of agreement

## 3) Plot agreement

```{r make-plot}
ba_plot(
  data = tbl,
  var1 = infrared,
  var2 = rectal,
  title = "Infrared vs rectal temperature",
  caption = "Lines show bias and limits of agreement with confidence bounds."
)
```

## Next steps

After this baseline workflow, explore:

1. transformed analyses (`transform = "log"` or `"logit"`)
1. grouped analyses via `group =` in `ba_stat()` and `ba_plot()`
