---
title: "Grouped and stratified Bland-Altman analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Grouped and stratified Bland-Altman analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why group the analysis?

Agreement can differ across biological or operational subgroups.  
`ggBA` supports grouped summaries and faceted plots to inspect those patterns.

```{r setup}
library(dplyr)
library(knitr)
library(tidyr)
library(ggBA)
```

## Build paired data once

```{r prep-data}
tbl <- temperature |>
  pivot_wider(names_from = method, values_from = temperature)
```

## Grouped statistics by treatment

```{r grouped-stats}
treatment_stats <- ba_stat(
  data = tbl,
  var1 = infrared,
  var2 = rectal,
  group = treatment
)

treatment_stats |>
  filter(parameter %in% c("bias", "lloa", "uloa")) |>
  tidyr::pivot_wider(names_from = parameter, values_from = value) |>
  arrange(treatment) |>
  kable(digits = 3)
```

This table highlights subgroup-level shifts in bias and spread.

## Faceted Bland-Altman plots

```{r grouped-plot}
ba_plot(
  data = tbl,
  var1 = infrared,
  var2 = rectal,
  group = treatment,
  colour = visit,
  title = "Agreement by treatment group",
  caption = "Each panel has treatment-specific Bland-Altman summary lines."
)
```

## Optional: log-scale grouped analysis

If multiplicative differences are more relevant, add `transform = "log"`:

```{r grouped-log}
ba_stat(
  data = tbl,
  var1 = infrared,
  var2 = rectal,
  group = treatment,
  transform = "log"
) |>
  filter(parameter == "bias") |>
  mutate(geom_ratio = exp(value)) |>
  select(treatment, geom_ratio) |>
  arrange(treatment) |>
  kable(digits = 3)
```
